Wall-to-wall mapping of peat depth from Lidar terrain and airborne radiometrics in Norwegian landscapes

### 2024-10-07 ms/ms.Rmd
ms: DTW preds

>>>>

The last terrain-based predictor we included was the depth-to-water index [@murphyMappingWetlandsComparison2007].
This index approximates a location's vertical height above the surface water feature that it is likely to drain towards.
It is calculated as the minimum cumulative slope (scaled by cell size) to a surface water feature [eq. 5 in @murphyTopographicModellingSoil2009].
We calculated unitless slope from the \unit{1\,m} DTM using the Whitebox software.
Also using Whitebox, we defined surface water features from the DTM by filling depressions and then calculating flow accumulation to define catchment areas for each cell [@schonauerSpatiotemporalPredictionSoil2021; @schonauerRcodeCalculatingDepthtowater2021].
This catchments area layer was then thresholded at seven different levels (*flow initiation areas* from \unit{0.25\,ha} to \unit{16\,ha}) to estimate surface water features under moisture scenarios varying from wet to dry [@murphyModellingMappingTopographic2011; @agrenEvaluatingDigitalTerrain2014; @schonauerSpatiotemporalPredictionSoil2021].
In addition, all surface water features mapped in the AR5 dataset were also transferred to the raster layer.
For each of the seven surface water layers, we derived the depth-to-water index using the *Distance Accumulation* tool in ArcGIS Pro (v.3.1, ESRI, USA), which has an efficient algorithm to find the cumulative distance over a cost surface to the least-cost source.

### 2024-10-07 ms/ms.tex
ms: DTW preds

\caption{\label{tab:preds}Quantitative predictors of peat depth.}
\centering
\begin{tabular}[t]{l|l}
\hline
name & description\\
\hline
radK & Potassium ground concentration\\
\hline
radTh & Thorium ground concentration\\
\hline
radU & Uranium ground concentration\\
\hline
radTC & Total count of gamma radiation\\
\hline
elevation & Mean elevation\\
\hline
slope1m & Mean of 1 m slope\\
\hline
TPI1m & Mean of 1 m topographic position index\\
\hline
TRI1m & Mean of 1 m terrain ruggedness index\\
\hline
roughness1m & Mean of 1 m roughness\\
\hline
slope10m & 10 m slope\\
\hline
TPI10m & 10 m topographic position index\\
\hline
TRI10m & 10 m terrain ruggedness index\\
\hline
roughness10m & 10 m roughness\\
\hline
MRVBF & Multi-resolution valley bottom flatness\\
\hline
TWI5m & Mean of 5 m topographic wetness index\\
\hline
TWI10m & 10 m topographic wetness index\\
\hline
TWI20m & Bilinear interpolation of 20 m topographic wetness index\\
\hline
TWI50m & Bilinear interpolation of 50 m topographic wetness index\\
\hline
DTW2500 & Depth-to-water index, flow initiation area of 2500 m\textasciicircum{}2\\
\hline
DTW5000 & Depth-to-water index, flow initiation area of 5000 m\textasciicircum{}2\\
\hline
DTW10000 & Depth-to-water index, flow initiation area of 10000 m\textasciicircum{}2\\
\hline
DTW20000 & Depth-to-water index, flow initiation area of 20000 m\textasciicircum{}2\\
\hline
DTW40000 & Depth-to-water index, flow initiation area of 40000 m\textasciicircum{}2\\
\hline
DTW80000 & Depth-to-water index, flow initiation area of 80000 m\textasciicircum{}2\\
\hline
DTW160000 & Depth-to-water index, flow initiation area of 160000 m\textasciicircum{}2\\
\hline
\end{tabular}
\end{table}

@@

>>>>

\caption{\label{tab:preds}Quantitative predictors of peat depth.}
\centering
\begin{tabular}[t]{l|l}
\hline
name & description\\
\hline
radK & Potassium ground concentration\\
\hline
radTh & Thorium ground concentration\\
\hline
radU & Uranium ground concentration\\
\hline
radTC & Total count of gamma radiation\\
\hline
elevation & Mean elevation\\
\hline
slope1m & Mean of 1 m slope\\
\hline
TPI1m & Mean of 1 m topographic position index\\
\hline
TRI1m & Mean of 1 m terrain ruggedness index\\
\hline
roughness1m & Mean of 1 m roughness\\
\hline
slope10m & 10 m slope\\
\hline
TPI10m & 10 m topographic position index\\
\hline
TRI10m & 10 m terrain ruggedness index\\
\hline
roughness10m & 10 m roughness\\
\hline
MRVBF & Multi-resolution valley bottom flatness\\
\hline
TWI5m & Mean of 5 m topographic wetness index\\
\hline
TWI10m & 10 m topographic wetness index\\
\hline
TWI20m & Bilinear interpolation of 20 m topographic wetness index\\
\hline
TWI50m & Bilinear interpolation of 50 m topographic wetness index\\
\hline
DTW2500 & Depth-to-water index, flow initiation area of 0.25 ha\\
\hline
DTW5000 & Depth-to-water index, flow initiation area of 0.5 ha\\
\hline
DTW10000 & Depth-to-water index, flow initiation area of 1 ha\\
\hline
DTW20000 & Depth-to-water index, flow initiation area of 2 ha\\
\hline
DTW40000 & Depth-to-water index, flow initiation area of 4 ha\\
\hline
DTW80000 & Depth-to-water index, flow initiation area of 8 ha\\
\hline
DTW160000 & Depth-to-water index, flow initiation area of 16 ha\\
\hline
\end{tabular}
\end{table}

@@

The last terrain-based predictor we included was the depth-to-water index \citep{murphyMappingWetlandsComparison2007}.
This index approximates a location's vertical height above the surface water feature that it is likely to drain towards.
It is calculated as the minimum cumulative slope (scaled by cell size) to a surface water feature \citep[eq. 5 in][]{murphyTopographicModellingSoil2009}.
We calculated unitless slope from the \unit{1\,m} DTM using the Whitebox software.
Also using Whitebox, we defined surface water features from the DTM by filling depressions and then calculating flow accumulation to define catchment areas for each cell \citep{schonauerSpatiotemporalPredictionSoil2021, schonauerRcodeCalculatingDepthtowater2021}.
This catchments area layer was then thresholded at seven different levels (\emph{flow initiation areas} from \unit{0.25\,ha} to \unit{16\,ha}) to estimate surface water features under moisture scenarios varying from wet to dry \citep{murphyModellingMappingTopographic2011, agrenEvaluatingDigitalTerrain2014, schonauerSpatiotemporalPredictionSoil2021}.
In addition, all surface water features mapped in the AR5 dataset were also transferred to the raster layer.
For each of the seven surface water layers, we derived the depth-to-water index using the \emph{Distance Accumulation} tool in ArcGIS Pro (v.3.1, ESRI, USA), which has an efficient algorithm to find the cumulative distance over a cost surface to the least-cost source.
